---
title: "<span style = 'color: grey;'> Bound Constrained Optimal Design of Multilevel Regression Discontinuity Designs and Randomized Controlled Trials </span>"
author: "Metin Bulus & Nianbo Dong"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{COSA Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cosa)
```

To install and load the package:
```{r eval = FALSE}
install.packages("cosa")
library(cosa)
```

The following examples demonstrate how to perform bound constrained optimal sample size allocation (BCOSSA) for a two-level cluster randomized trial (CRT) and for the corresponding cluster-level regression discontinuity (CRD) design under three primary constraints. Note: `NULL` arguments are provided for clarity, otherwise they do not have to be explicit.

## Primary Constraint on the Total Cost

```{r, message=FALSE, fig.width=7, fig.height=5, warning=FALSE}
# cost constrained - optimize p and n2
# CRT('order = 0' or 'rhots = 0')
crt <- cosa.crd2(order = 0,
                 constrain = "cost", cost = 12500, 
                 cn1 = c(5, 2), cn2 = c(50, 20),
                 es = .20,  power = .80, rho2 = .20,
                 g2 = 1, r21 = .20, r22 = .30, 
                 p = NULL, n1 = 20,  n2 = NULL)
# comparisons to CRDs
# CRD w/ linear score variable interacting with the treatment
crd1 <- cosa.crd2(order = 1, interaction = TRUE,
                  constrain = "cost", cost = 12500,
                  cn1 = c(5, 2), cn2 = c(50, 20),
                  es = .20,  power = .80, rho2 = .20,
                  g2 = 1, r21 = .20, r22 = .30,
                  p = .386, n1 = 24,  n2 = NULL)
# CRD w/ quadratic score variable interacting with the treatment
crd2 <- cosa.crd2(order = 2, interaction = TRUE,
                  constrain = "cost", cost = 12500,
                  cn1 = c(5, 2), cn2 = c(50, 20),
                  es = .20,  power = .80, rho2 = .20,
                  g2 = 1, r21 = .20, r22 = .30,
                  p = .386, n1 = 24,  n2 = NULL)

# example plots
par(mfrow = c(2, 3), mai = c(.6, .6, .6, .2))
# compare minimum detectable effect size and 95% CI
plot(crt, ypar = "mdes", xpar = "n2",
     ylim = c(.10, .90), xlim = c(10, 800), 
     ylab = "MDES (with Power = .80)", xlab = "Number of Clusters",
     main = expression(CRT), locate = TRUE)
plot(crd1, ypar = "mdes", xpar = "n2",
     ylim = c(.10, .90), xlim = c(10, 800),
     ylab = "MDES (with Power = .80)", xlab = "Number of Clusters",
     main = expression(CRD (S + TS)), locate = TRUE)
plot(crd2, ypar = "mdes", xpar = "n2",
     ylim = c(.10, .90), xlim = c(10, 800),
     ylab = "MDES (with Power = .80)", xlab = "Number of Clusters",
     main = expression(CRD (S + S^2 + TS + TS^2)), locate = TRUE)
# compare statistical power 
plot(crt, ypar = "power", xpar = "n2",
     ylim = c(.10, .85), xlim = c(10, 800), 
     ylab = "Power (for ES = .20)", xlab = "Number of Clusters",
     main = expression(CRT), locate = TRUE)
plot(crd1, ypar = "power", xpar = "n2",
     ylim = c(.10, .85), xlim = c(10, 800), 
     ylab = "Power (for ES = .20)", xlab = "Number of Clusters",
     main = expression(CRD(S + TS)), locate = TRUE)
plot(crd2, ypar = "power", xpar = "n2",
     ylim = c(.10, .85), xlim = c(10, 800), 
     ylab = "Power (for ES = .20)", xlab = "Number of Clusters",
     main = expression(CRD (S + S^2 + TS + TS^2)), locate = TRUE)

```

As seen from the MDES and power plots, a little less than 150 clusters are needed to obtain the benchmark power rate of 80% for the CRT (more than 200 clusters for the CRD). Precise number of clusters can be found via placing the primary constraint on either the effect size or power rate. 

## Primary Constraint on the Effect Size

```{r, message=FALSE, warning=FALSE}
# cost constrained - optimize p and n2
# CRT('order = 0' or 'rhots = 0')
cosa.crd2(order = 0,
          constrain = "es", es = .20,  
          cn1 = c(5, 2), cn2 = c(50, 20),
          power = .80, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30, 
          p = NULL, n1 = 20,  n2 = NULL)
# comparisons to CRDs
# CRD w/ linear score variable interacting with the treatment
cosa.crd2(order = 1, interaction = TRUE,
          constrain = "es", es = .20,  
          cn1 = c(5, 2), cn2 = c(50, 20),
          power = .80, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30,
          p = .389, n1 = 24,  n2 = NULL)
# CRD w/ quadratic score variable interacting with the treatment
cosa.crd2(order = 2, interaction = TRUE,
          constrain = "es", es = .20,  
          cn1 = c(5, 2), cn2 = c(50, 20),
          power = .80, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30,
          p = .389, n1 = 24,  n2 = NULL)

```

## Primary Constraint on the Power Rate

```{r, message=FALSE, warning=FALSE}
# cost constrained - optimize p and n2
# CRT('order = 0' or 'rhots = 0')
cosa.crd2(order = 0,
          constrain = "power", power = .80, 
          cn1 = c(5, 2), cn2 = c(50, 20),
          es = .20, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30, 
          p = NULL, n1 = 20,  n2 = NULL)
# comparisons to CRDs
# CRD w/ linear score variable interacting with the treatment
cosa.crd2(order = 1, interaction = TRUE,
          constrain = "power", power = .80, 
          cn1 = c(5, 2), cn2 = c(50, 20),
          es = .20, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30,
          p = .389, n1 = 24,  n2 = NULL)
# CRD w/ quadratic score variable interacting with the treatment
cosa.crd2(order = 2, interaction = TRUE,
          constrain = "power", power = .80,  
          cn1 = c(5, 2), cn2 = c(50, 20),
          es = .20, rho2 = .20,
          g2 = 1, r21 = .20, r22 = .30,
          p = .389, n1 = 24,  n2 = NULL)

```
--o--
